REPLACE ME

### 2023-08-14 Markdown/Manuscript/Child_RMDs/MS_Body.Rmd
circulated draft
stuck
Redo with accontable, reusable and updatable themes
up to fw
updated intro

@@

@@

>>>>

```{r child="../Child_RMDs/XX_Funs.Rmd", echo=FALSE}
```
```{r child="../Child_RMDs/MS_Tables.Rmd", echo=FALSE}
```
```{r child="../Child_RMDs/MS_Figures.Rmd", echo=FALSE}
```
<!-- Add child RMDs for results content -->

@@

# Conclusion

@@

## Acknowledgement {-}
The authors would like to acknowledge the contribution of John Gillam who provided advisory input to this research.

### 2023-08-14 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
updated intro

**Transparency** has been recommended as a core criterion for assessing ethical public health modelling practice [@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommended that model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Notably, the same guidelines, published over ten years ago, did not include recommendations on sharing model code and data. However, more recent guidance recommends publicly dissemination of healthcare model artefacts using online repository services [@Erdemir2020]. Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. An essential component of quality assuring health economic models is verification - ensuring that calculations are correct and consistent with model specifications [@techver2019]. The extensiveness of verification checks in models implemented as software projects can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

@@

Model **reusability** is the second attribute that we identified as being core to ethical CHEM implementations. Making a CHEM’s code, data and documentation publicly available is now considered “gold-standard” practice but is not sufficient to promote model re-use. Computational implementations that store model algorithms and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) facilitate selective modification of model components. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task), has been recommended as good practice for scientific computing [@Wilson_2017] and can facilitate model generalisability. The concept of inheritance [@8717448], a feature of object-oriented programming, facilitates selective editing of algorithms when transferring models to different decision contexts [@ready4oop2022]. This benefit can be further enhanced if model developers adopt a modular approach, in which a model is constructed from multiple reusable and replaceable sub-models (modules) [@pan2021modular]. To grant permissions to others to use, test and adapt models and their components, health economists can avail of two broad categories of open source licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
The **reusability** of a CHEM implementation can be assessed against the criteria:

-	R1: Model code and data are implemented to facilitate both generalizability and transferability.

@@

To remain valid for longer, models should be continually updated and refined as new evidence emerges and healthcare systems evolve [@Jenkins2021; @@garcia2021cost]. Ensuring that a model is regular reviewed to identify and implement required improvements is a receommended defence against model validity decay [@calder2018computational]. Key enablers of sustainable maintenance of open source research software are committed, adequately resourced core development team and active user community [@info:doi/10.2196/20028]. Currently, the core development team for a CHEM will be typically be funded to produce a project end-point deliverable whose specifications are well defined early in the project. For more complex and multi-purpose CHEMs, particularly those designed to be incorporated into decision support systems, it may be better for development teams to adopt Agile Software Development [@beck2001manifesto], an approach that has been recommended for complex public health software projects [@10.3389/fpubh.2022.899874]. An Agile model will be less clearly specified in the initial project plan, but will instead continually develop in response to the requirements and feedback of users, who are provided with an initial, simplified working version of the model at the earliest feasible opportunity. Online communities can be an efficient means of engaging model users in testing each version of a model, identifying issues and suggesting improvements. Services such as GitHub [@github2007] provide collaborative code development tools [@MERGEL2015464] that help elicit, integrate and reconcile contributions from multiple contributors and to ensure each update is uniquely identifiable and retrievable. It is important that verification checks are rerun with each model update, a task that can be automated using the software development practice of Continuous Integration [@CI2017].  The risk of model revisions having unintended consequences for third party users can be mitigated through the use of deprecation conventions [@zhou2016api] that take an informative and staged approach to retiring outdated model code and data.

Assessment criteria for an updatable CHEM:

-	U1: Resources and infrastructure are in place to support sustained development, testing, maintenance and version control of a model in collaboration with model users.

>>>>

**Transparency** has been  as a core identified as a core feature of ethical public health modelling practice [@10.3389/fpubh.2017.00068]. Guidance on transparency in health economic modelling recommended that model code and data should be clearly documented, potentially with different versions for technical and non-technical users [@Eddy2012]. Notably, the same guidelines, published over ten years ago, did not include recommendations on sharing model code and data. However, more recent guidance recommends publicly dissemination of healthcare model artefacts using online repository services [@Erdemir2020]. Repositories such as Zenodo [@Zenodo2013] and Dataverse [@Dataverse2007] provide persistent storage solutions that generate a Digital Object Identifier (DOI) for each code and data collection. An essential component of quality assuring health economic models is verification - ensuring that calculations are correct and consistent with model specifications [@techver2019]. The extensiveness of verification checks in models implemented as software projects can be reported using the concept of code coverage [@ERICWONG2010188] - the proportion of model code that has been explicitly tested. Tests should ideally combine both unit tests (to verify that small, isolated sections of code produce the correct output when run independently) and acceptance tests (to verify that the correct output is produced when multiple code components are run together to perform tasks that meet core user-requirements [@martin2003agile]). The nature and extent of individual model authorship contributions can become unclear when models are implemented over longer time-frames with a large and changing group of collaborators [@thompson2022escape]. This issue can be addressed by use of online repository services such as GitHub [@github2007], that provide citation tools and can transparently record all individual code contributions to a modelling project over its lifecycle.

@@

Model **reusability** is the second attribute that we identify as being core to ethical CHEM implementations. Making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, yet is not sufficient to promote model re-use.  Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task), is good scientific computing practice [@Wilson_2017] and facilitates model reuse without adaptation (generalisability). To support reuse with adaptation (transferability), computational implementations can store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code).  Transferability can also be advanced through the use of object-oriented programming approaches [@ready4oop2022] that take advantage of the concept of inheritance [@8717448] to allow models to be first duplicated and then selectively modified. Generalisability and transferability are further enhanced if computational implementations adopt a modular approach, in which a model is constructed from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. To grant permissions to others to use, test and adapt models and their components, health economists can avail of two broad categories of open source licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
<!-- are relatively easier to selectively modify. -->
The **reusability** of a CHEM implementation can be assessed against the criteria:

-	R1: Model code and data are implemented to facilitate both generalisability and transferability.

@@

Models should be **updatable** so that they remain valid for longer, evolving as new evidence emerges and the systems being modelled change [@Jenkins2021; @@garcia2021cost]. Ensuring that a model is regularly reviewed to identify and implement required improvements is a recommended defense against model validity decay [@calder2018computational]. Sustainable maintenance of open source research software requires both a core development team and an active user community [@info:doi/10.2196/20028]. Online communities can be an efficient means of engaging model users in testing each version of a model, identifying issues and suggesting improvements. Services such as GitHub [@github2007] provide collaborative code development tools [@MERGEL2015464] that help elicit, integrate and reconcile contributions from multiple contributors and to ensure each update is uniquely identifiable and retrievable. It is also important that verification checks are rerun with each model update, a task that can be automated using the software development practice of Continuous Integration [@CI2017].  The risk of model revisions having unintended consequences for third party users can be mitigated through the use of deprecation conventions [@zhou2016api] that take an informative and staged approach to retiring outdated model code and data.
<!-- Currently, the core development team for a CHEM will be typically be funded to produce a project end-point deliverable whose specifications are well defined early in the project. For more complex and multi-purpose CHEMs, particularly those designed to be incorporated into decision support systems, it may be better for development teams to adopt Agile Software Development [@beck2001manifesto], an approach that has been recommended for complex public health software projects [@10.3389/fpubh.2022.899874]. An Agile model will be less clearly specified in the initial project plan, but will instead continually develop in response to the requirements and feedback of users, who are provided with an initial, simplified working version of the model at the earliest feasible opportunity. -->

The **updatability** of a CHEM implementation can be assessed against the criteria:

-	U1: Technical infrastructure is in place to support sustained development, testing, maintenance and version control of a model in collaboration with model users.

### 2023-08-14 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
updated intro

To support the collaborative development of CHEMs that meet TRU assessment criteria, we have created a prototype software framework called ready4. We have designed this framework as a lightweight extension to existing software development infrastructure that could be used to implement CHEMs that are both open-source (for transparency and reusability) and modular (for reusability and updatability).

## 	Framework libraries 
To work within the popular open-source programming environment R [@RCORE2022], the ready4 software framework is implemented as six development version R code libraries. The R libraries collectively provide model developers with tools to combine R's functional and object-oriented programming paradigms [@10.1214/13-STS452] in authoring CHEMs, supplying those CHEMs with data and using CHEMs to implement reproducible modelling analyses. The six novel libraries and the preexisting third-party R libraries they depend on are summarised in Table \@ref(tab:cpkgs).

Standardization and interoperability are core requirements of implementing a modular approach. Model modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering]. Each CHEM module authored with the ready4 framework will include both a data structure (specifying the required properties of data that can validly be supplied to a module) and a set of algorithms (specifying the operations that can be performed on data contained in a module instance). The foundational framework library, called ready4, defines a template module data structure (using R’s S4 class system) from which all CHEM module data structures will be created and a novel syntax that enable module algorithms to be consistently named. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs authored with the framework and for partially automating updates to a project documentation website.

>>>>

To support the development of CHEMs that meet TRU assessment criteria, we have created a prototype software framework called ready4. We have designed this framework to enable ethical computational implementations of health economic models that are both open-source (for transparency and reusability) and modular (for reusability and updatability). The framework aims to support the standardization and interoperability that are core enablers of implementing a modular and open source approaches. Model modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering], even when these modules are authored by different teams.

## 	Framework libraries 
To work within the popular open-source programming language R [@RCORE2022], the ready4 software framework is implemented as R code libraries.

The R language supports functional and object-oriented programming paradigms [@10.1214/13-STS452]. Both these paradigms are used in the ready4 framework to provide tools for authoring CHEMs, supplying those CHEMs with data and using CHEMs to implement reproducible modelling analyses. 
<!-- Each CHEM module authored with the ready4 framework will include both a data structure (specifying the required properties of data that can validly be supplied to a module) and a set of algorithms (specifying the operations that can be performed on data contained in a module instance). -->

As R is an open source language, each new R library typically depends on a number of other R libraries that have been written by third parties. As the number of third party dependencies of an R library grows so does the fragility of that library (e.g. the library may cease to work as intended due to changes in one of its dependency libraries). To reduce the fragility of our framework we implemented it as multiple R libraries rather than one R library. In total we authored six novel R libraries to implement the ready4 framework, all of which have distinct purposes and depend on have unique third-party R library dependencies (Table \@ref(tab:cpkgs)).
<!-- development version -->

The foundational framework library, called ready4, defines a template module data structure (using R’s S4 class system) from which all CHEM module data structures will be created and a novel syntax that enable module algorithms to be consistently named. The ready4 library also contains tools for retrieving web based information on model modules, datasets and analysis programs authored with the framework and for partially automating updates to a project documentation website.

### 2023-08-14 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
updated intro

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four].  In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. These representations can be described in words and pictures (a conceptual model), in equations (a mathematical model) or in computer code (a computational model). The predictions reported in health economic studies are typically generated by the execution of a computer program which applies a computational model to compatible data inputs (e.g., parameter values) and performs a sequence of numeric calculations.

Computational models are now widely used to inform health policy and system design [@dakin2015influence; @Erdemir2020; @LO2022100570; @Christene006827]. This level of influence has concomitant ethical responsibilities for model developers that are often poorly understood and inadequately fulfilled [@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational].

Computational health economic models (CHEMs) can be implemented using specialized commercial software or authored as bespoke software projects in a programming language. Advantage of commercial modelling tools are simplicity and ease of use, but a software development approach facilitates development of models that may be more realistic, transparent, reusable and adaptable [@incerti2019r]. For CHEMs that are implemented as software projects, a major early decision is selection of an appropriate software framework. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. However, software frameworks can be challenging and time consuming to create [@edwin2014software] and then difficult for others to learn, often requiring developers to undergo specialist training [@myllarniemi2018development]. There is also a risk that a software framework may become excessively complex over time [@edwin2014software].

We are developing a model to explore multiple economic questions relating to the mental health of young people aged 12 to 25.  When making choices about how to implement this model computationally, we wished to facilitate ethical development and use. However, we are not aware of any software framework for implementing CHEMs that adhere to explicitly stated ethical requirements.

>>>>

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four].  In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and computer code (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by applying the computational model. The conceptual, mathematical and computational representations of a model are assumed to be isomorphic. Independent assessment of the validity of this assumption depends in part on the approach taken to computational model implementation and dissemination.

<!-- the execution of a computer program that applies a computational model to compatible data inputs (e.g., parameter values) and performs a sequence of numeric calculations. -->
<!-- the predictions reported in health economic studies are now almost always -->

Computational health economic models (CHEMs) can be implemented using specialized commercial software or authored as software development projects in an open-source programming language such as Python [@python2009] or R [@RCORE2022]. Advantages of commercial modelling tools include simplicity and ease of use, but a software development approach may facilitate computational models that are more transparent, reusable and adaptable [@incerti2019r]. The most appropriate approach for a specific CHEM project may depend on both the desired features of the model being developed and the resource and time constraints within which the project needs to be delivered (see Table \@ref(tab:proscons)). Although health economists report strong interest in making CHEMs more transparent and accessible [@Pouwels2022], open-source CHEMs remain rare [@Feenstra2022; @Emerson2019].

```{r, child = child_docs_ls$one, echo=FALSE}
```
<!-- on considerations such as model comprehensiveness, generalisability, reproducibility and transferability -->
A major early decision in open-source CHEM projects is the selection of an appropriate software framework. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources, e.g., code and data files, that can be modified to suit specific needs. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.

Advantages of using software frameworks include facilitating code reuse and extension, promoting good programming practice and the capability to provide enhanced functionality and performance without additional effort by developers [@edwin2014software]. However, software frameworks can be challenging and time consuming to create [@edwin2014software] and then difficult for others to learn, often requiring model developers to undergo specialist training [@myllarniemi2018development]. There is also a risk that a software framework may become excessively complex over time [@edwin2014software].

<!-- We are developing a model to explore multiple economic questions relating to the mental health of young people aged 12 to 25.   -->
<!-- When making choices about how to implement this health economic model computationally, we wished to facilitate ethical development and use. -->
A high-level software framework using R has previously been developed with the primary aim of improving the transparency of CHEMs [@alarid2019need]. Novel software frameworks may be necessary to facilitate CHEM implementations that meet broader ethical criteria. The ethical responsibilities of computational modellers are often poorly understood and inadequately fulfilled [@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational]. Guidance on developing, selecting and using software frameworks for ethical CHEM implementations is scarce.

<!-- Transparency has been proposed as one of the core ethical responsibilities of public health computational modellers [@10.3389/fpubh.2017.00068].  -->

<!-- Although computational models are widely used to inform health policy and system design [@dakin2015influence; @Erdemir2020; @LO2022100570; @Christene006827], -->

### 2023-08-14 Markdown/Manuscript/Child_RMDs/MS_Results.Rmd
working my way through supervisors comments
updated intro

```{r child="../Child_RMDs/XX_Funs.Rmd", echo=FALSE}
```
```{r child="../Child_RMDs/MS_Tables.Rmd", echo=FALSE}
```
```{r child="../Child_RMDs/MS_Figures.Rmd", echo=FALSE}
```
<!-- Add child RMDs for results content -->

## Ethical requirements and assessment criteria
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

>>>>

## Assessing ethical computational implementations
```{r child="../Child_RMDs/MS_Imp_Guidelines.Rmd", echo=FALSE}
```

### 2023-08-14 Markdown/Manuscript/Parent_PDF/Parent_PDF.Rmd
updated intro

---
bibliography: ../Data/references.bib
output: 
  bookdown::pdf_book: 
    base_format: rticles::arxiv_article
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
  bookdown::pdf_document2: 
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
    toc: no
## CUSTOM_AUTHORS_PLACEHOLDER ##
params:
  X: NULL
abstract: "TO BE REPLACED"
# abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"
graphics: yes
csl: plos.csl
---

>>>>

---
bibliography: ../Data/references.bib
output: 
  bookdown::pdf_book: 
    base_format: rticles::springer_article # rticles::arxiv_article
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
  # bookdown::pdf_document2: 
  #   includes:
  #     in_header: preamble.tex
  #   pandoc_args: 
  #     - "--metadata-file=../Header/header_common.yaml"
  #     - '--lua-filter=../Filters/scholarly-metadata.lua'
  #     - '--lua-filter=../Filters/author-info-blocks.lua'
  #   toc: no
  #   keep_tex: true

params:
  X: NULL
abstract: "TO BE REPLACED"
# abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"
graphics: yes
#csl: plos.csl
title: REPLACE ME
titlerunning: TEMPORARY
authorrunning: Hamilton, Gao, Wiesner et al.
thanks: | 
    The study was funded by ADD DETAILS HERE. 
keywords: one, two, three, four, five, six
#PACS: 
#- PAC1
#- superPAC

# MSC:
# - MSC code 1
# - MSC code 2
classoptions: 
  - sn-basic      
  # - Numbered      # Optional: Use numbered references instead of namedate references (only for sn-nature, sn-basic, sn-vancouver, sn-chicago, sn-mathphys or sn-nature)
  # - referee       # Optional: Use double line spacing 
  # - lineno        # Optional: Add line numbers
  # - iicol         # Optional: Double column layour
#biblio-style: spbasic
# bibstyle options spbasic(default), spphys, spmpsci
# header-includes:
# - \usepackage{booktabs}
# - \usepackage{longtable}
# - \usepackage{array}
# - \usepackage{multirow}
# - \usepackage{wrapfig}
# - \usepackage{float}
# - \usepackage{colortbl}
# - \usepackage{pdflscape}
# - \usepackage{tabu}
# - \usepackage{threeparttable}
# - \usepackage[normalem]{ulem}
# - \usepackage{threeparttablex}
# - \usepackage{makecell}
---
